Classless Java

### 2016-03-23 OOPSLA16/sections/Introduction.tex
related work

One important advantage of the use of abstract state operations and
automatic type-refinement is that it offers a new form of
\emph{type-safe covariant mutable state}. That is, in IB programming,
it is possible to refine \emph{mutable} ``fields'' in subtypes. This is
typically forbidden in CB programming languages (including Java and
Scala), since it is widely known that \emph{naive} type-refinement of
mutable fields is not type-safe. While covariant refinement of mutable
fields is supported by some type systems~\cite{bruce98astatically,McDirmid01jiazzi,ernst01family,ernst06virtual}, this requires
significant complexity and restrictions to ensure that all uses of
covariant state are indeed type-safe.

>>>>

One important advantage of the use of abstract state operations and
automatic type-refinement is that it offers a new form of
\emph{type-safe covariant mutable state}. That is, in IB programming,
it is possible to refine \emph{mutable} ``fields'' in subtypes. This is
typically forbidden in CB programming languages (including Java and
Scala), since it is widely known that \emph{naive} type-refinement of
mutable fields is not type-safe. While covariant refinement of mutable
fields is supported by some type systems~\cite{bruce98astatically,bruce1994paradigmatic,ernst06virtual,Saito2013933}, this requires
significant complexity and restrictions to ensure that all uses of
covariant state are indeed type-safe.
